---
title       : Coursera DataProducts Assignment
subtitle    : Decision Quality Shiny App
author      : Joseph Kahn
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---



## Goals

Before begining a data science task, it is most important to determine the purpose of the analysis.
Often, we are performing analysis in support (directly or indirectly) of decision making.
A broader picture of decision making considers the entire decision analysis cycle.

Based on the [Decision Education Foundation](http://www.decisioneducation.org), there are 6 elements to acheiving decision quality:

1. Appopriate Frame
2. Clear Objectives
3. Creative Alternatives
4. Useful Information
5. Sound Reasoning
6. Commitment to Action

This [Decision Quality Shiny App](http://josephmkahn.shinyapps.io/DecisionQuality) helps to facilitate good decision making by highlighting the key steps, and 
allowing users to visually display the weak points in overall decision quality.

---

## Decision Quality Scale 

Each element of decision quality can be rated from 0% up to 100%. Although we could spend forever 
on any element, we mark 100% as the point at which it is no longer worth the extra effort to improve that element.
The table below gives guidance at what 0% and 100% should look like for each element.

```{r, results='asis', echo=FALSE, fig.height=5}

library(fmsb)
library(xtable)

all.min <- 0
all.max <- 100

maxmin <- data.frame(
  Frame=c(all.max, all.min)
  , Commitment=c(all.max, all.min)
  , Logic=c(all.max, all.min)
  , Information=c(all.max, all.min)
  , Alternatives=c(all.max, all.min)
  , Objectives=c(all.max, all.min)
)



GuideDF <- data.frame(
  At_0_Pct = c("Divergent understanding of purpose, but plowing ahead anyway",
               "Missing key stakeholders, unclear on prime value drivers, ignoring important intangibles",
               "One option / Business as Usual",
               "Lacking info on key inputs, point estimates rather than ranges/distributions",
               "Missing key issues, while over-complex on others. Results are confusing to the decision maker",
               "Lack of stakeholder support, organizational hurdles, continual rework, too early or too late"),
  
  At_100_Pct = c("Alignment by key stakeholders and experts, and committed to the process",
                 "Quantitative tradeoffs among stakeholders' key metrics, including time and risk preferences",
                 "Multiple compelling Strategies",
                 "Relevant info assessed and reviewed from trusted sources. Ranges and uncertainties well characterized",
                 "Best choice(s) are clear, model is trusted, insights verified and robust",
                 "Stakeholders bought in, team aligned, resources committed"))

rownames(GuideDF) = c("Appropriate Frame",
                      "Clear Objectives",
                      "Creative Alternatives",
                      "Useful Information",
                      "Sound Reasoning",
                      "Commitment to Action")

colnames(GuideDF) = c("At 0%","At 100%")

print(xtable(GuideDF),type='html')

```

---

## Example: lack of creative alternatives 

This plot demonstrates where we have not considered any alternatives other than business as usual. As a result, stakeholders have not committed to the strategy.


```{r, echo=FALSE, fig.height=5}


    dat <- data.frame(
      Frame=85
      , Commitment=20
      , Logic=90
      , Information=85
      , Alternatives=20
      , Objectives=95
    )
    
    dat <- rbind(maxmin,dat)
    
    
    radarchart(dat
               ## ## , axistype=0
               , seg=5
               , plty=1
               , vlabels=c("Appropriate Frame"
                           , "Commitment to\nAction"
                           , "Sound\nReasoning"
                           , "Useful Information"
                           , "Creative\nAlternatives"
                           , "Clear Objectives"
               )
               , pcol="deeppink4"               #blue
               , plwd=3
               , cglwd=2
               , pfcol=c("darksalmon", 1, NA, NA, NA, NA)
               , cex=1.5
               , cex.lab=1.5
               , cex.main=1.5
               , cex.sub=1.5
               , pdensity=20
    )
    



```


---

## Example: High decision quality

This plot demonstrates where all elements of decision quality have been adequately addressed, helping stakeholders to commit to action.


```{r, echo=FALSE, fig.height=5 }


    dat <- data.frame(
      Frame=95
      , Commitment=100
      , Logic=90
      , Information=90
      , Alternatives=90
      , Objectives=95
    )
    
    dat <- rbind(maxmin,dat)
    
    
    radarchart(dat
               ## ## , axistype=0
               , seg=5
               , plty=1
               , vlabels=c("Appropriate Frame"
                           , "Commitment to\nAction"
                           , "Sound\nReasoning"
                           , "Useful Information"
                           , "Creative\nAlternatives"
                           , "Clear Objectives"
               )
               , pcol="deeppink4"               #blue
               , plwd=3
               , cglwd=2
               , pfcol=c("darksalmon", 1, NA, NA, NA, NA)
               , cex=1.5
               , cex.lab=1.5
               , cex.main=1.5
               , cex.sub=1.5
               , pdensity=20
    )
    



```


---


## Conculsions

This Decision Quality Shiny App assists users in thinking about the elements of decision quality
The App is live at Shiny.io [http://josephmkahn.shinyapps.io/DecisionQuality](http://josephmkahn.shinyapps.io/DecisionQuality).
Code and this presentation is archived on github [https://github.com/josephmkahn/DataProductsAssignment](https://github.com/josephmkahn/DataProductsAssignment) in the ShinyApp folder.

Acknowledgements to:

* The Decision Education Foundation [http://www.decisioneducation.org](http://www.decisioneducation.org)
* fmsb R package, for the "radar" plot
* Vineet Joshi, for assisting with cration of radar plots
* Coursera for hosting the class
* Rstudio for creating Shiny and hosting shiny.io
* Creators of knitr and slidify
* Our data products and data science instructors at Johns Hopkins!


